Leadership across cultures and contexts: an exploratory analysis

### 2020-06-17 leadership_across_cultures_contexts.Rmd
Revision 2 edits to clarify and reduce word count

---
title: "Leadership across cultures and contexts: an exploratory analysis"
author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
date: "4/25/2019"
output: 
  bookdown::html_document2:
    number_sections: true
bibliography: Leader2.bib
csl: apa-6th-edition.csl
---

@@

* Leaders receive import material, social, and reproductive benefits

@@

>It's good to be king, if just for a while. - Tom Petty

@@

To operationalize leader qualities and functions from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records. KS then coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered. After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars). KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.

In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori* based on theoretical literature. For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories. See Table \@ref(tab:costbenefitvars). Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny). The distinction is that costs and benefits were transactional whereas qualities were descriptive (e.g., leaders obtain wives vs. have wives).

@@

ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*). We illustrate our coding with an example from Andamanese hunter-gatherers [@az02-001, p. 45]:

Each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy. We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category. We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.

Each document was coded for publication date and the ethnographer(s) sex [some metadata collected/confirmed with @mehr_natural_2016]. For additional details on ethnographic sampling and creation of the text record database, see @garfield_evolutionary_2019.

@@

Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis. See the supplementary information (SI) for brief descriptions of these methods. Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex, using logistic mixed effects models fit with the lme4 package. Our sixth goal was to evaluate the features in light of influential evolutionary theories of leadership. Drawing on the empirically derived features as a guide, we aimed to construct "elements" of leadership that corresponded to elements of influential theories, examining intersections of these elements across all text records.

@@

The analyses computing levels of evidence for each dimension used the entire set of text records. Analyses investigating lower dimensional representations used a smaller set that omitted rows with no evidence for any dimension (i.e., all zeros), and dimensions with very little evidence (almost all zeros). All analyses were conducted with `r R.version.string`.

@@

```{r geographic, fig.width=15, fig.cap="The geographical distributions of cultures in this analysis. Symbol shapes indicate subsistence strategy and symbol size indicates the number of documents available for that culture."}
library(maps)
world_map <- map_data("world")
p_map <- ggplot(world_map, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), fill="lightgray", colour = "white") +
  geom_point(
    data = leader_cult, 
    aes(
      c_longitude, 
      c_latitude, 
      colour = subsistence, 
      shape = subsistence, 
      size = documents
    )
  ) +
  coord_fixed() +
  labs(x='', y='', shape = 'Subsistence', colour = 'Subsistence', size = 'Documents') +
  guides(shape = guide_legend(override.aes = list(size = 4))) +
  theme_bw(15)
ggsave(filename = 'Figures/map.pdf', p_map)
p_map
```

@@

## Distribution of text records by subsistence strategy and group type

Subsistence strategy was confounded with continental region. African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Figure S\@ref(fig:regionsubsistence)).

Subsistence strategy and group structure were also confounded. Hunter-gatherer text records often referred to residential leadership (`r hg_residential_pct`% of the total sample of hunter-gatherer text records) and kin leadership (`r hg_kin_pct`%), but rarely referred to supracommunity leadership (`r hg_supra_pct`%). Agricultural text records, on the other hand, rarely referred to residential leadership (`r agri_residential_pct`%) but often referred to supracommunity leadership (`r agri_supra_pct`%). Across all subsistence types, political groups at the community level were the most frequent context, with the exception of agriculturalists where supracommunity political groups were the most frequent. Leadership beyond the community level was relatively rare outside of agricultural populations and was especially rare among hunter-gatherers. See Figure \@ref(fig:grouptype).

@@

For the number of text records for each continental region, subsistence strategy and group context, see Table S\@ref(tab:fourfactorsumm).

## Leader sex

@@

## Possible sources of bias

@@

We investigated if there were biases in the discovery of evidence for any of our `r length(all_study_vars)` leader dimensions by publication date of source ethnographic texts, female gender of any author or co-author, or total pages of ethnography from each culture. Publication dates ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure S\@ref(fig:doc-years)). There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors). Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author. Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`, and this was weakly correlated with the number of text records on leadership retrieved for a particular culture (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`). See Figure S\@ref(fig:eHRAFpages).

We fit `r length(all_study_vars)` logistic mixed effects models with each binary leadership dimension as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures. We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995]. After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but recent publications were significantly more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*). See Figure S\@ref(fig:pubdate) and SI for details.

We then tested if text records with a female author or co-author were more likely to discuss female leaders. Of the `r sum(tab_coauthor['FALSE',])` text records with no female author or co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership. Of the `r sum(tab_coauthor['TRUE',])` text records with a female author or co-author, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders. After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`). Because we found minimal evidence of bias, we do not control for any bias variables in subsequent analyses.

@@

For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (as described in the Methods section). See Figure \@ref(fig:varsupport). For a table of all coefficients and 95% CIs, see Table S\@ref(tab:glmmcoefs).

@@

Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1 (on the scale of the linear predictor). See Figures S\@ref(fig:FunctionRE) and S\@ref(fig:QualityRE). In a few cases, one or the other was exactly 0, which can occur in models like ours that have many random effect levels (e.g., `r length(unique(all_data$author_ID))` authors) and a relatively small data set [@bates2015].

@@

We investigated four possibles sources of variation in evidence for our `r length(all_study_vars)` dimensions of leadership: group context, subsistence strategy, continental region, and leader sex. Specifically, we fit separate multiple logistic mixed effects regression models of each leadership dimension as a function of these four factors, with random intercepts for author nested within culture. We then compared the AIC value of each model to a model with only the fixed and random intercepts. Models whose AIC values were at least two less than the intercept-only models were deemed to vary by those factors [@burnham2002].

@@

These factors were confounded, i.e., group context was associated with subsistence strategy, which was associated with continental region. See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence). Specifically, evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017]. Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories. These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among nonindustrial populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010]. The only example of supracommunity leadership among hunter-gatherers was from the Andamanese by @az02-001[p. 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section). One interpretation of the negative relationship between evidence for political leadership at the community level or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.

@@

@@

The 60-culture Probability Sample Files of the HRAF developed by @naroll_proposed_1967 is based on culturally and geographically diverse populations which are well covered in the ethnographic record. The distribution of cultures, however, is not equal across categories of subsistence strategies; agriculturalists and horticulturalists are relatively over-represented compared to hunter-gatherers and pastoralists. The ethnographic sample of text developed in the *leadershipdata* package is also not equally balanced across all subsistence strategy categories. See Figure S\@ref(fig:subsisttree).

@@

The eHRAF provides information on the time frame of field work associated with each document. For some documents the span of associated fieldwork is given for the exact months and years the ethnographer was in the field. For most documents one or more spans of years are given. Figure \@ref(fig:field-years) plots the earliest year and the latest year for which field work occurred for each document. The time spans plotted do not depict the full span of fieldwork necessarily, but simply illustrate the time span during which fieldwork occurred. We omitted data for one document from North America with field work date ranges prior to 1805. For all documents, the average earliest year in which field work occurred was `r signif(mean(documents$d_field_date_start, na.rm=T),4)` and the average latest years in which field work occurred was `r signif(mean(documents$d_field_date_end, na.rm=T),4)`.

@@

@@

@@

```{r heatmapregion, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each continental region. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}
ggsave(filename = 'Figures/heatmap_region.pdf', p_heatmap_region + theme_minimal(15))
p_heatmap_region + theme_minimal(15)
```

@@

Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA]. NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$. NMF estimates $W$ and $H$ by minimizing a measure of distance between $X$ and $WH$ [we used the default "brunet" algorithm, which uses the Kullback-Leibler divergence; @gaujoux2010].

>>>>

---
title: "Leadership across cultures and contexts: an exploratory analysis"
author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
date: "4/25/2019"
output: 
  bookdown::word_document2:
    number_sections: true
bibliography: Leader2.bib
csl: apa-6th-edition.csl
---

@@

* Leaders receive important material, social, and reproductive benefits

@@

<!-- It's good to be king, if just for a while. - Tom Petty -->

@@

To operationalize leader qualities and functions from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records. KS then coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered. After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table S\@ref(tab:qualfuncvars). KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.

In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori* based on theoretical literature. For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories. See Table S\@ref(tab:costbenefitvars). Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny). The distinction is that costs and benefits were transactional whereas qualities were descriptive (e.g., leaders obtain wives vs. have wives).

@@

ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*). For an illustration of the coding process, see the *Coding example* section of the SI.

Each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy.

@@

Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, minimal spanning trees and nearest neighbors (mst-kNN), and hierarchical cluster analysis. See the supplementary information (SI) for brief descriptions of these methods. Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex, using logistic mixed effects models fit with the lme4 package. Our sixth goal was to evaluate the features in light of influential evolutionary theories of leadership. Drawing on the empirically derived features as a guide, we aimed to construct "elements" of leadership that corresponded to elements of influential theories, examining intersections of these elements across all text records.

@@

The analyses computing levels of evidence for each dimension used the entire set of text records. Analyses investigating lower dimensional representations used a smaller set that omitted rows with no evidence for any dimension (i.e., all zeros), and dimensions with very little evidence (almost all zeros).

Throughout, we compare regression models of the same outcome variable using the Akaike information criterion (AIC). $AIC=2k - 2 ln(\hat{L})$, where $k$ is the number of model parameters and $\hat{L}$ is the maximum likelihood. AIC penalizes models with more parameters and rewards models with better fit. Hence, a model with a lower AIC value is considered to be superior.

All analyses were conducted with `r R.version.string`.

@@

```{r geographic, fig.width=15, fig.cap="The geographical distributions of cultures in this analysis. Symbol shapes indicate subsistence strategy and symbol size indicates the number of documents available for that culture. Axes are degrees latitude and longitude."}
library(maps)
world_map <- map_data("world")
p_map <- ggplot(world_map, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), fill="lightgray", colour = "white") +
  geom_point(
    data = leader_cult, 
    aes(
      c_longitude, 
      c_latitude, 
      colour = subsistence, 
      shape = subsistence, 
      size = documents
    )
  ) +
  coord_fixed() +
  labs(x='', y='', shape = 'Subsistence', colour = 'Subsistence', size = 'Documents') +
  guides(shape = guide_legend(override.aes = list(size = 4))) +
  theme_bw(15)
ggsave(filename = 'Figures/map.pdf', p_map)
p_map
```

@@

Subsistence strategy, group structure and continental region were confounded. Hunter-gatherer text records often referred to residential and kin leadership, but rarely to supracommunity leadership, whereas agricultural text records rarely referred to residential leadership but often referred to supracommunity leadership. Community level political groups were frequent across all subsistence types. African text records were predominantly from horticultural societies whereas North American text records were predominantly from hunter-gatherer societies. See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence), and Table S\@ref(tab:fourfactorsumm).

@@

@@

@@

We investigated if there were biases in the discovery of evidence for any of our `r length(all_study_vars)` leader dimensions by publication date of source ethnographic texts, female gender of any author or co-author, or total pages of ethnography from each culture. Because we found minimal evidence of bias, we do not control for any bias variables in subsequent analyses. For details, see the *Bias* section of the SI.

@@

For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (see Tables S\@ref(tab:qualfuncvars) and S\@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (as described in the Methods section). See Figure \@ref(fig:varsupport). For a table of all coefficients and 95% CIs, see Table S\@ref(tab:glmmcoefs).

@@

The random intercepts indicated that support for many variables did vary modestly by ethnographer and culture. Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1 (on the scale of the linear predictor). See Figures S\@ref(fig:FunctionRE) and S\@ref(fig:QualityRE). In a few cases, one or the other was exactly 0, which can occur in models like ours that have many random effect levels (e.g., `r length(unique(all_data$author_ID))` authors) and a relatively small data set [@bates2015].

@@

We investigated four possibles sources of variation in evidence for our `r length(all_study_vars)` dimensions of leadership: group context, subsistence strategy, continental region, and leader sex. Specifically, we fit separate multiple logistic mixed effects regression models of each leadership dimension as a function of these four factors, with random intercepts for author nested within culture. We then compared the AIC value of each model to a model with only the fixed and random intercepts. Models whose AIC values were at least two less than the intercept-only models were deemed to outperform the intercept-only model [@burnham2002].

@@

These factors were confounded, i.e., group context was associated with subsistence strategy, which was associated with continental region. See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence). Specifically, evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017]. Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories. These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among nonindustrial populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010]. The only example of supracommunity leadership among hunter-gatherers was from the Andamanese by @az02-001[p. 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in *Coding example* section of SI). One interpretation of the negative relationship between evidence for political leadership at the community level or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.

@@

## Operationalization of Leader Qualities and Functions variables

## Operationalization of Leader and Follower Costs and Benefits variables

@@

The 60-culture Probability Sample Files of the HRAF developed by @naroll_proposed_1967 is based on culturally and geographically diverse populations which are well covered in the ethnographic record. We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category. We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.

The distribution of cultures, however, is not equal across categories of subsistence strategies; agriculturalists and horticulturalists are relatively over-represented compared to hunter-gatherers and pastoralists. The ethnographic sample of text developed in the *leadershipdata* package is also not equally balanced across all subsistence strategy categories. See Figure S\@ref(fig:subsisttree).

@@

Each document was coded for publication date and the ethnographer(s) sex [some metadata collected/confirmed with @mehr_natural_2016]. The eHRAF provides information on the time frame of field work associated with each document. For some documents the span of associated fieldwork is given for the exact months and years the ethnographer was in the field. For most documents one or more spans of years are given. Figure \@ref(fig:field-years) plots the earliest year and the latest year for which field work occurred for each document. The time spans plotted do not depict the full span of fieldwork necessarily, but simply illustrate the time span during which fieldwork occurred. We omitted data for one document from North America with field work date ranges prior to 1805. For all documents, the average earliest year in which field work occurred was `r signif(mean(documents$d_field_date_start, na.rm=T),4)` and the average latest years in which field work occurred was `r signif(mean(documents$d_field_date_end, na.rm=T),4)`.

@@

For additional details on ethnographic sampling and creation of the text record database, see @garfield_evolutionary_2019.

## Coding example

We illustrate our coding with an example from Andamanese hunter-gatherers [@az02-001, p. 45]:

@@

Publication dates ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure S\@ref(fig:doc-years)). There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors). Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author. Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`, and this was weakly correlated with the number of text records on leadership retrieved for a particular culture (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`). See Figure S\@ref(fig:eHRAFpages).

We fit `r length(all_study_vars)` logistic mixed effects models with each binary leadership dimension as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures. We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995]. After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but recent publications were significantly more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*). See Figure S\@ref(fig:pubdate).

We then tested if text records with a female author or co-author were more likely to discuss female leaders. Of the `r sum(tab_coauthor['FALSE',])` text records with no female author or co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership. Of the `r sum(tab_coauthor['TRUE',])` text records with a female author or co-author, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders. After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).

@@

```{r heatmapregion, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each continental region. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}
ggsave(filename = 'Figures/heatmap_region.pdf', p_heatmap_region + theme_minimal(15), width=12)
p_heatmap_region + theme_minimal(15)
```

@@

Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA]. A minimum spanning tree (MST) takes a distance matrix of the column vectors (computed here using binary distance) and then finds a graph that connects all nodes with the shortest total distance. MST was combined with k-nearest neighbor (kNN), where a complete graph is formed of each node's k nearest neighbors. The result is the intersection of these two graphs [@inostroza2008].

<!-- NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$. NMF estimates $W$ and $H$ by minimizing a measure of distance between $X$ and $WH$ [we used the default "brunet" algorithm, which uses the Kullback-Leibler divergence; @gaujoux2010]. -->
